Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning

### 2013-10-30 04_experiment.tex
Sec 4-6 proof

We recruited 1032 US Participants via Amazon Mechanical Turk, an online crowd-sourcing tool. Each participant played a one-shot referential game either as a speaker or as a listener and received a payment of 5 cents. If a participant contributed multiple trials, only the first was included. All participants passed the attention check described below.

@@

Each participant saw a context consisting of three objects (Figure \ref{exp_context} as an example) and reported on the number of objects in the context as an attention check. Participants in the listener conditions were told to imagine someone was talking to them and used some utterance (depending on the exact condition) to refer to exactly one of the objects in the context. Then they were asked to choose what they thought to be the intended referent. Participants in the speaker conditions were told to imagine they were talking to someone and they wanted to refer to one of the objects (indicated by a small red arrow\footnote{Participants were told that the arrow was only for illustration and thus the person they were talking to could not see it.}). Then they were asked to choose between two words to refer to it.

@@

In order to minimize the effect of the confounding factors, we counterbalanced the positions and colors of the objects, as well as the orders of the candidate words in the speaker conditions. Thus the unique color in a context could be either blue (as in Fig. \ref{exp_context}) or green and the order of the objects could be any of the permutations\footnote{Due to the scale of our preliminary experiment we did not counterbalance the shape of the objects, i.e., a context always has one square and two circles. }.

@@

There are four major designs in our own experiment that differ from the original experiment in \cite{Frank}. First of all, the original experiment used the betting paradigm, i.e., participants were asked to bet over a range of options and they were instructed that the amount of money that is bet on an option should reflect their confidence that the option is correct. Even though the betting paradigm has the merit of providing us with graded reponses from each individual participant, the caveat is that it is unclear whether it measures the belief or the action. This can lead to confusion when we are to fit the model predictions to the empirical data without knowing whether they are directly comparable. In addition, since the betting paradigm is more or less introspective in nature, it tends to be not very accurate. Thus we used forced choice in our experiment instead, which clearly measures the action and does not rely on introspection.
Secondly, since we decided to investigate the influence of the speaker's preference as well as the listener's perceptual salience, we focused on contexts equivalent to Figure \ref{context}, which are most typical in Gricean accounts of scalar implicature, and examined different features separately.
Thirdly, we only included the features of color and shape to ensure the vocabulary to be common knowledge, excluding the feature of texture. Finally, we did not use dotted line surrounding an object as the way to indicate the target to the speaker, as it could emphasize the feature of shape and thus add confound in the experiment. Instead we used a small arrow pointing to the target.

@@

\

The results of the speaker and listener conditions of our experiment are shown as follows.

@@

\begin{table}[htb]   
  \caption{Speaker Conditions}
  \centering
  \begin{tabular}{c|ccccc}
   Target  & ``Green'' & ``Square'' & ``Blue'' & ``Circle'' & Total\\ 
     \hline
\textcolor{green!65}{\Large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- &144\\
\textcolor{blue!65}{\Huge{$\bullet$}}           & --        &   --      & $119$ & $25$ & 144\\
\textcolor{green!65}{\Huge{$\bullet$}}          & $63$        &   --    &  --   & $81$             &144
  \end{tabular}

@@

From the data in Table \ref{table:speaker}, we can see that the speakers tend to choose the optimal feature more often when the target has the unique shape than when it has the unique color ($\chi^2=8.5, p<.01$). Even though they seem to prefer the feature of shape when the target's both features are shared, the difference is not statistically significant from uniform random choice ($\chi^2=2.25, p=.13$).

@@

\begin{table}[htb] 
  \caption{Listener Conditions}
  \centering
\begin{tabular}{c|cccc}
   & \quad \textcolor{green!65}{\Large{$\blacksquare$}}&  \textcolor{green!65}{\Huge{$\bullet$}}& \textcolor{blue!65}{\Huge{$\bullet$}} & Total\\ 
     \hline
 Salience   \quad  & \quad  $71$      \quad    &   $30$    \quad     & $139$ \quad     & $240$      \\
 ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & $0$ \quad     & $180$     \\
 ``Circle''   &  \quad $1$        &    $62$        & $117$ \quad     & $180$
  \end{tabular}

@@

The result of the salience condition will be used as the empirical estimation of contextual salience. For the other two conditions, the object with both features shared (the green circle in the above table, but note that we counterbalanced colors) is what the Gricean pragmatic account predicts to be the target. The result in Table \ref{table:listener} shows that when the message is ``Green'', listeners prefer the green circle which is predicted by the Gricean pragmatics, while they tend to stick to the blue circle which is more perceptually salient when the message is ``Circle''. The behavioral patterns in both conditions are significantly different from uniform random choice, and they significantly differ from each other in respect of whether they conform to the predictions by Gricean pragmatic theory.

>>>>

We recruited 1032 US participants via Amazon's Mechanical Turk, an
online crowd-sourcing tool. Each participant played a one-shot
referential game either as a speaker or as a listener and received a
payment of 5 cents. If a participant contributed multiple trials, only
the first was included. All participants passed the attention check
described below.

@@

Each participant saw a context consisting of three objects (Figure
\ref{exp_context} as an example) and reported on the number of objects
in the context as an attention check. Participants in the listener
conditions were told to imagine someone was talking to them who used
some utterance (depending on the specific condition tested) to refer
to exactly one of the objects in the context. Then they were asked to
choose what they thought to be the intended referent. Participants in
the speaker conditions were told to imagine they were talking to
someone and they wanted to refer to one of the objects (indicated by a
small red arrow\footnote{Participants were told that the arrow was
  only for illustration and thus the person they were talking to could
  not see it.}). Then they were asked to choose between two words to
refer to it.

@@

In order to minimize the effect of the confounding factors, we
counterbalanced the positions and colors of the objects, as well as
the orders of the candidate words in the speaker conditions. Thus the
unique color in a context could be either blue (as in
Fig. \ref{exp_context}) or green and the order of the objects could be
any of the permutations.\footnote{Due to the scale of our preliminary
  experiment we did not counterbalance the shape of the objects, i.e.,
  a context always had one square and two circles.}

@@

There are four major design choices in our own experiment that differ
from the original experiment in \cite{Frank}. First of all, the
original experiment used a \emph{betting paradigm}, i.e., participants
were asked to bet over a range of options and they were instructed
that the amount of money that is bet on an option should reflect their
confidence that the option is correct. Even though the betting
paradigm has the merit of providing us with graded reponses from each
individual participant, the caveat is that it is unclear whether it
measures beliefs or actions. This can lead to confusion when we are to
fit the model predictions to the empirical data without knowing
whether they are directly comparable. In addition, since the betting
paradigm is more or less introspective in nature, it tends to be not
very accurate. Thus we used a \emph{forced choice design} instead,
which clearly measures the action, does not rely on introspection and,
technically speaking, provides for a straightforward likelihood
function to be used in subsequent statistical model comparison.
Secondly, since we decided to investigate the influence of the
speaker's preference as well as the listener's perceptual salience, we
focused on contexts equivalent to Figure \ref{context}, which require
reasoning that is highly reminiscent of Gricean accounts of scalar
implicature calculation, and examined different features separately.
Thirdly, we only included the features of color and shape to ensure
the vocabulary to be common knowledge, excluding the feature of
texture. Finally, we did not use dotted lines surrounding an object as
the way to indicate the target to the speaker, as \cite{Frank} did,
because that might unduly emphasize the feature of shape and thus be a
confound in the experiment. Instead we used a small arrow pointing to
the target.

@@

The results of the speaker and listener conditions of our experiment
are as follows.

@@

\begin{table}[htb]   
  \caption{Speaker Conditions}
  \centering
  \begin{tabular}{c|ccccc}
   Target  & ``Green'' & ``Square'' & ``Blue'' & ``Circle'' & Total\\ 
     \hline
\textcolor{green!65}{\large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- &144\\
\textcolor{blue!65}{\Huge{$\bullet$}}           & --        &   --      & $119$ & $25$ & 144\\
\textcolor{green!65}{\Huge{$\bullet$}}          & $63$        &   --    &  --   & $81$             &144
  \end{tabular}

@@

From the data in Table \ref{table:speaker}, we can see that the
speakers tend to choose the optimal feature more often when the target
has the unique shape than when it has the unique color ($\chi^2=8.5,
p<.01$). Even though they seem to prefer the feature of shape when
both of the target's features are shared, the difference is not
statistically significant from uniform random choice ($\chi^2=2.25,
p=.13$).

@@

\begin{table}[htb] 
  \caption{Listener Conditions}
  \centering
\begin{tabular}{c|cccc}
   & \quad \textcolor{green!65}{\Large{$\blacksquare$}}&  \textcolor{green!65}{\Huge{$\bullet$}}& \textcolor{blue!65}{\Huge{$\bullet$}} & Total\\ 
     \hline
 Salience   \quad  & \quad  $71$      \quad    &   $30$    \quad     & $139$ \quad     & $240$      \\
 ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & -- \quad     & $180$     \\
 ``Circle''   &  \quad $1$        &    $62$        & $117$ \quad     & $180$
  \end{tabular}

@@

The result of the salience condition will be used as the empirical
estimation of contextual salience. For the other two conditions, the
object with both features shared (the green circle in the above table,
but note that we counterbalanced colors) is what the Gricean pragmatic
account predicts to be the target. The result in Table
\ref{table:listener} shows that when the message is ``Green'',
listeners prefer the green circle which is predicted by the Gricean
pragmatics, while they tend to stick to the blue circle which is more
perceptually salient when the message is ``Circle''. The behavioral
patterns in both conditions are significantly different from uniform
random choice, and they significantly differ from each other in
whether they conform to the predictions by Gricean pragmatic theory.

### 2013-10-30 06_conclusion.tex
Sec 4-6 proof

Our experiment and model comparison suggest the following: (1) On the conceptual level, 
it is helpful to clarify the distinction between belief, goal and action, and the Bayesian framework provides us with a natural way to structure these components into a formal model. By examining each component systematically, we can explicitly spell out the underlying assumptions and formulate alternative hypotheses to be tested. (2) On the empirical level, we tested the RSA model with a family of variants motivated from the game-theoretic pragmatic perspectives on communication, and in particular, we investigated the speaker's preference as well as the listener's perceptual salience. Our data showed a more intricate picture about the roles various factors play in pragmatic reasoning about referential expressions --- what is commonly known is the Gricean pragmatic components, while either side have his own private bias unknown to the other. Also, an action-based view might better reflect the goal of communication, at least in the forced-choice experimental setting.

Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to a better understanding of the phenomenon and further improvement in the formal model's predictive power. Of course, more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.

>>>>

Our experiment and model comparison suggest the following: (1) On the
conceptual level, it is helpful to clarify the distinction between
belief, goal and action, and the Bayesian framework provides us with a
natural way to structure these components into a formal model. By
examining each component systematically, we can explicitly spell out
the underlying assumptions and formulate alternative hypotheses to be
tested. (2) On the empirical level, we tested the RSA model with a
family of variants motivated from a game theoretic perspective on
communication, and in particular, we investigated the speaker's
preference as well as the listener's perceptual salience. Our data
showed a more intricate picture of the roles that various factors play
in pragmatic reasoning about referential expressions. We found
evidence for a correlation between the speaker's and the listener's
rationality. Either side appears to have his own biases, but appears
to be negligent of the other's. Also, an action-based view might
better reflect the goal of communication, at least in the
forced-choice experimental setting. Understanding the relation between
the forced-choice and the betting paradigms would be an important next
step to gain more insight into the differences between action- and
belief-based goal structures.

Our work shows that careful conceptual analysis of the design choices
for quantitative models can lead to a better understanding of the
phenomenon and further improvement in the formal model's predictive
power. Of course, our restricted empirical data can only serve as a
start and more data is needed to fuel further model comparison towards
a more robust pragmatic theory of people's probabilistic reasoning
about referential expressions.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% TeX-PDF-mode: t
%%% End:
